---
title: "hitting"
author: "Trey Chase"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
hitters <- read_csv("hitters.csv")
```

```{r}
dim(hitters)
head(hitters)
```

```{r}
library(tidyverse)

sum(colSums(is.na(hitters)))
#observed 35 NAs in the dataset
#30 blast speed, 5 exit velo

#dropping na values
hitters = hitters |>
  drop_na()
```

```{r}
library(ggplot2)

ggplot(hitters, aes(x=exit_velo_mph_x)) + 
  geom_histogram(color = "black", fill = "orange") +
  labs(
    title = "Distribution of Exit Velocities"
  )
```

### Principal Component Analysis

```{r}
train_df = hitters |>
  #removing non-numeric/relevant columns
  select(
    -session,
    -session_swing
  )

predictors = train_df |>
  select(
    -exit_velo_mph_x
  )

response = train_df |>
  select(
    exit_velo_mph_x
  )
```

```{r}
library(pls)

pcomps = prcomp(predictors, scale = TRUE)
par(mar = c(5, 5, 5, 5), cex = 0.2)  
biplot(pcomps, scale = 0, cex = 1.5)
```

```{r}
pcomps$rotation
```

```{r}
pca_var <- pcomps$sdev^2
pca_var
```

```{r}
pve <- pca_var / sum(pca_var)
pve
```

```{r}
par(mfrow = c(1, 2))
plot(pve, xlab = "Principal Component",
ylab = "Proportion of Variance Explained", ylim = c(0, 1),
type = "b")
plot(cumsum(pve), xlab = "Principal Component",
ylab = "Cumulative Proportion of Variance Explained",
ylim = c(0, 1), type = "b")
```

```{r}
pcr_m = pcr(exit_velo_mph_x~.,
             data=train_df,
             scale = TRUE,
             validation = "CV")

summary(pcr_m)
validationplot(pcr_m, val.type = "MSEP")
```

```{r}
pca_fit <- prcomp(predictors, scale. = TRUE)


top_components <- data.frame(pca_fit$x[, 1:20], exit_velo_mph_x = response)


pcr_fit <- lm(exit_velo_mph_x ~ ., data = top_components)


summary(pcr_fit)
```

### Partial Least Squares

```{r}
pls_fit <- plsr(exit_velo_mph_x ~ ., data = train_df, scale = TRUE, validation = "CV")
summary(pls_fit)
validationplot(pls_fit, val.type = "MSEP")
```

```{r}
train_df <- na.omit(train_df)


pls_fit <- plsr(exit_velo_mph_x ~ ., data = train_df, scale = TRUE, ncomp = 5)


top_pls_components <- data.frame(pls_fit$scores[, 1:5], exit_velo_mph_x = train_df$exit_velo_mph_x)


lm_fit <- lm(exit_velo_mph_x ~ ., data = top_pls_components)


summary(lm_fit)
```

```{r}
loadings <- pls_fit$loadings[, 1:5]

top_contributors <- list()

for (i in 1:5) {
  component_loadings <- loadings[, i]
  max_variable <- rownames(loadings)[which.max(abs(component_loadings))]
  top_contributors[[paste0("Component_", i)]] <- max_variable
  cat("Top contributing variable for Component", i, ":", max_variable, "\n")
}

print(top_contributors)

```
* 'bat_speed_xy_max_x' : maximum resultant angular velocity of the bat - just x and y components (deg/sec) 
* 'pelvis_angle_fp_x' : pelvis angle (x) at foot plant - using K-Vest conventions (deg)
* 'hand_speed_mag_max_x' : maximum resultant hand speed between first move and contact (deg/sec) 
* 'pelvis_angle_hs_x' : pelvis angle (x) at heel strike - using K-Vest conventions (deg)
* 'torso_angle_hs_z' : torso angle (x) at heel strike - using K-Vest conventions (deg)

