Author: Trey Chase
Date: sys.date()


```{r loading packages, echo=FALSE, warning= FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(GGally)
library(tidyr)
library(stats)
library(rsample)
```


```{r loading data}
df <- read_csv("/Users/treychase/Downloads/poi_metrics.csv")
```
```{r df structure}
dim(df)
head(df)
summary(df)
```

```{r data selection}
#obtaining all quantitative variables, dropping all rows with na
df = df |>
  select_if(
    is.numeric, 
  ) |>
  drop_na() |>
  select(
    -(session) #removing session column
  )
```


```{r correlation plot, echo=FALSE}
#checking multi collinearity condition, as well as obtaining the most correlated value with pitch speed
library(ggcorrplot)
corr_matrix = cor(df)
diag(corr_matrix) = NA

first_column <- corr_matrix[,1] #looking at the pitch speed column
max_value <- max(first_column, na.rm = TRUE) #finding the maximum correlated variable with pitch speed
max_var <- rownames(corr_matrix)[which(first_column == max_value)]


max_value
max_var
```

```{r correlation plot}
cor_plot = ggcorrplot(corr_matrix) 
ggsave("dl_corr_plot.png", plot = cor_plot, width = 30, height = 30)
```
Saving the correlation plot to examine which variables are correlated with pitch speed.



```{r splitting data}
set.seed(123)  # For reproducibility
split <- initial_split(df, prop = 0.8) #splitting data into test and training data
train_df <- training(split)
test_df <- testing(split)
```



```{r elbow transfer linear plot}
ggplot(train_df, aes(x=elbow_transfer_fp_br, y=pitch_speed_mph)) + 
  geom_point() +
  geom_smooth(method = "lm", color="blue") +
  labs(
    x = "Elbow Transfer FP",
    y= "Pitch Speed MPH",
    title = "Elbow Transfer Force vs. Velocity"
  )
```

```{r elbow transfer plotting transformed}
ggplot(train_df, aes(x=(elbow_transfer_fp_br), y=log(pitch_speed_mph))) + 
  geom_point() +
  geom_smooth(method = "lm", color="blue") +
  labs(
    x = "Elbow Transfer Force Plate",
    y= "Pitch Speed MPH",
    title = "Elbow Transfer Force vs. Log Velocity"
  )
```






```{r transformation plot}
model = lm(pitch_speed_mph ~ elbow_transfer_fp_br, data=train_df)
summary(model)
plot(model)
```


```{r cooks distance}
#removing the high leverage values using Cook's distance
cooks_dist <- cooks.distance(model)
cooks_threshold <- 4 / nrow(train_df)
high_cooks <- which(cooks_dist > cooks_threshold)
train_df <- train_df[-high_cooks, ]
```


```{r plotting new model fit}
ggplot(train_df, aes(x=log(elbow_transfer_fp_br), y=log(pitch_speed_mph))) + 
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2),color="blue") +
  labs(
    x = "Log Elbow Transfer Force Plate",
    y= "Log Pitch Speed MPH",
    title = "Elbow Transfer Force Squared vs. Log Velocity"
  )
```

The polynomial looks like a better fit than the line.

```{r fitting a polynomial model}
poly_model = lm(log(pitch_speed_mph) ~ poly(log(elbow_transfer_fp_br), 2), data = train_df)
summary(poly_model)
plot(poly_model)
```


```{r shoulder transfer model}
shoulder_model = lm(log(pitch_speed_mph) ~ poly(shoulder_transfer_fp_br, 2), data = train_df)
summary(shoulder_model)
plot(shoulder_model)
```

```{r removing high leverage points}
cooks_dist <- cooks.distance(shoulder_model)
cooks_threshold <- 4 / nrow(train_df)
high_cooks <- which(cooks_dist > cooks_threshold)
train_df <- train_df[-high_cooks, ]
```

The model yield an R-Squared of about 0.5 after removing outliers. This is okay for non-linear regression, but there is room for improvement.


```{r fitting models}
thorax_model = lm(log(pitch_speed_mph) ~ poly((log(thorax_distal_transfer_fp_br)),2), data = train_df)
summary(thorax_model)
```

We obtain an R^2 of roughly 0.42.

```{r plot of three predcitors}
ggplot(train_df, aes(x=(thorax_distal_transfer_fp_br), y=log(pitch_speed_mph))) + 
  geom_point() +
  geom_smooth(method = "lm",formula = y ~ poly(x,2), color="blue") +
  labs(
    x = "Thorax Distal Transfer Force Plate",
    y= "Log Pitch Speed MPH",
    title = "Thorax Distal Transfer Force Squared vs. Log Velocity"
  ) #thorax

ggplot(train_df, aes(x=(shoulder_transfer_fp_br), y=log(pitch_speed_mph))) + 
  geom_point() +
  geom_smooth(method = "lm",formula = y ~ poly(x,2), color="blue") +
  labs(
    x = "Shoulder Transfer Force Plate",
    y= "Log Pitch Speed MPH",
    title = "Shoulder Transfer Force Squared vs. Log Velocity"
  )

ggplot(train_df, aes(x=log(elbow_transfer_fp_br), y=log(pitch_speed_mph))) + 
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2),color="blue") +
  labs(
    x = "Log Elbow Transfer Force Plate",
    y= "Log Pitch Speed MPH",
    title = "Elbow Transfer Force Squared vs. Log Velocity"
  )
```


```{r multivariate model}
m_model = lm(log(pitch_speed_mph) ~ poly((log(thorax_distal_transfer_fp_br)),2) + poly(log(shoulder_transfer_fp_br, 2)) + poly(log(elbow_transfer_fp_br), 2), data = train_df)

summary(m_model)
plot(m_model)
```

Integrating the three most correlated variables, we get a multiple R^2 of 0.5 and an adjusted R^2 of 0.49. This is okay, but there could be a better model out there.


```{additive model}
gam(formula = pitch_speed_mph ~ s(elbow_absorption_fp_br))
```





